---
title: "Week 3 Demonstration"
format: html
editor: visual
---

## Set up

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
library(slider)
```

## Computing classical decomposition

Compute a classical decomposition for the following time series without using the 
`classical_decomposition()` function.

```{r}
aus_arrivals_jap <- aus_arrivals |>
  filter(Origin == "Japan") |>
  select(Quarter, Arrivals)
```

## Classical vs STL decomposition

Start with the following code snippet creating a time series of passengers 
flying on Ansett Airlines. Perform classical and STL decompositions and comment
on their differences. Which do you trust more?

```{r}
melsyd_economy <- ansett |>
  filter(Airports == "MEL-SYD", Class == "Economy") |>
  mutate(Passengers = Passengers/1000)
autoplot(melsyd_economy, Passengers) +
  labs(y = "Passengers ('000)")
```

## White noise and ACF plots

Create a time series of length 200 comprising i.i.d. standard Gaussian measurements.
Make a time plot of the time series as well as an ACF plot.
Compare this with the remainder term from an STL decomposition of `aus_arrivals_jap`.
What do you observe?

## Summary statistics

Glance through the summary statistics described in [Chapter 4.4](https://otexts.com/fpp3/other-features.html) of Hyndman and Athanasopoulos (2021).
What do the following summary statistics compute?

- Strength of trend
- Strength of seasonality
- `shift_level_max`
- `shift_level_index`
- `var_tiled_var`

When would they be useful?

## EDA with summary statistics

The follow code snippet generates summary statistics for the various time series in the `tourism` dataset.

```{r}
#| eval: FALSE
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))
```

Make a scatter plot of seasonal strength against trend strength and color the points according to `Purpose`.
What can you tell from the plot?

Perform PCA on the `tourism_features` dataframe and make a scatter plot of the first two PCs.
What do you observe?

## Time series clustering using summary statistics

Use summary statistics to create a model for classifying the time series in this [dataset](https://kdd.ics.uci.edu/databases/synthetic_control/synthetic_control.html).