---
title: "Week 4 Demonstration"
format: html
editor: visual
  markdown: 
    wrap: 72
---

## Set up

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
library(slider)
library(gridExtra)
library(broom)
```

## 1. Simple forecasting strategies and statistical models

### Mean method

Show that the mean method is derived from the statistical model shown in the video, i.e. $$
x_t = \theta + \epsilon_t
$$ with $\epsilon_1,\epsilon_2,\ldots,\epsilon_n \sim_{i.i.d.} \mathcal{N}(0,\sigma^2)$.

### Naive method

Show that the naive method is derived from the statistical model shown in the video, i.e. $$
x_t = x_{t-1} + \epsilon_t
$$ with $\epsilon_1,\epsilon_2,\ldots,\epsilon_n \sim_{i.i.d.} \mathcal{N}(0,\sigma^2)$.

### Seasonal naive method

What is the statistical model associated with the seasonal naive method?

### Linear trend method

What is the statistical model for the linear trend method?

### Drift method

Consider the statistical model $$
x_t = \theta + x_{t-1} + \epsilon_t.
$$ This is a random walk with drift $\theta$. What is the conditional distribution $p_\theta(x_t|x_{t-1})$? What is maximum likelihood estimate for $\theta$? If we use this for forecasting, what is the formula for the $h$-step ahead forecast?

## 2. Drift vs linear trend method

The drift method can be fit using the code `RW(y ~ drift())`. Starting with the following code snippet, compute forecasts using the drift method and the linear trend method for the population of Australia for the next 50 years .

```{r}
global_economy |> filter(Country == "Australia")
```

Which forecast looks better? Which prediction intervals are more realistic?

Can you explain the connection between this and the statistical modeling assumptions?

## 3. Seasonal naive method with drift

We have learnt one way to model both linear trend and seasonality by using time series decomposition. Here, we introduce another method.

Think of a modification of the seasonal naive model to incorporate drift. What is the statistical model? What is the forecast equation?

The method can be fit using the code `SNAIVE(y ~ drift())`.

## 4. Forecasting

Which of `NAIVE`, `SNAIVE`, `NAIVE(y ~ drift())` and `SNAIVE(y ~ drift())` are most appropriate for the following datasets?

-   Bricks (`aus_production`)

-   Household wealth (`hh_budget`)

## 5. Prediction intervals

Consider the `aus_arrivals` dataset. Filter the time series of arrivals from Japan to before 1995, and fit `NAIVE`, `SNAIVE`, `NAIVE(y ~ drift())` and `SNAIVE(y ~ drift())` . Use the fitted models to forecast the rest of the time series. Do their prediction intervals contain the truth?
