---
title: "Week 6 Demonstration"
format: html
editor: visual
---

## Set up

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
library(slider)
library(gridExtra)
```

## 1. ACVF, ACF, and stationarity

Consider the time series defined as $$
X_t = U_1 \sin(2\pi \omega t) + U_2\cos(2\pi\omega t),
$$ where $U_1$ and $U_2$ are independent random variables with zero mean and variance 1.

a.  Compute the ACVF of $(X_t)$.
b.  Is $(X_t)$ weakly stationary?
c.  Is $(X_t)$ strictly stationary? If not, what conditions can we put on $U_1$ and $U_2$ to make it stationary?

## 2. ACF vs sample ACF

Consider the linear trend model $$
X_t = \beta_0 + \beta_1 t + W_t.
$$

a.  Compute the ACF of $(X_t)$.
b.  Simulate a time series drawn from this model and plot its sample ACF.
c.  Why does the sample ACF not look like the population ACF function?
d.  Why does the asymptotic normality theorem for the ACF not apply?

## 3. Forecasting

Let $(X_t)$ be a zero mean stationary process with $\rho_X(h) \neq 0$ for some $h$

a.  Find $A$ minimizing $\mathbb{E}\lbrace (X_{n+h} - A X_n)^2 \rbrace$.
b.  Use this to make a forecast for $X_{n+h}$ given $X_1,\ldots,X_n$.
c.  What is the MSE for the forecast?
d.  What is the consequence of this for ACF values?

## 4. Residual analysis

a.  What is the difference between innovation residuals and regular residuals? Why do we perform residual analysis on the former rather than the latter?
b.  Starting with the following code snippet, fit several models to the Australian takeaway turnover time series and analyze their residuals. Which models have a good fit? Compare this with their CV error.

```{r}
takeaway <- aus_retail |>
  filter(Industry == "Takeaway food services") |>
  summarise(Turnover = sum(Turnover))
```
