---
title: "Week 2 Demonstration"
format: html
editor: visual
---

## Set up

We always load the following packages.

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
```

## Lags and differencing

Let's consider the `aus_arrivals` dataset and try to plot lagged arrival values.

```{r}
aus_arrivals |>
  mutate(LaggedArrivals = lag(Arrivals)) |>
  autoplot(LaggedArrivals)
```

Something seems a bit odd about this plot, especially the UK value. 
What is happening is that the lag function has treated the entire column as a 
single time series even though it contains 4 separate time series. 
To avoid this, we need to do a `group_by()` before applying mutate

```{r}
aus_arrivals |>
  group_by(Origin) |>
  mutate(LaggedArrivals = lag(Arrivals)) |>
  ungroup() |>
  autoplot(LaggedArrivals)
```

## Moving averages

Let's focus on the arrivals from Japan time series and compute moving averages.

```{r}
library(slider)

aus_jap_arrivals <- aus_arrivals |> 
  filter(Origin == "Japan") |>
  mutate(MA2 = slide_dbl(Arrivals, mean, .before = 1, .after = 0),
         MA2_alt = (Arrivals + lag(Arrivals)) / 2)
aus_jap_arrivals
```

Manually computing the moving average shows that the formula is correct. 
We now compute moving averages with a few different window sizes to see the 
effect of varying the window size.

```{r}
library(slider)
library(viridis)

aus_jap_arrivals <- aus_arrivals |> 
  filter(Origin == "Japan") |>
  mutate(ArrivalsMA2 = slide_dbl(Arrivals, mean, .before = 1, .after = 0),
         ArrivalsMA4 = slide_dbl(Arrivals, mean, .before = 2, .after = 1),
         ArrivalsMA6 = slide_dbl(Arrivals, mean, .before = 3, .after = 2),
         ArrivalsMA8 = slide_dbl(Arrivals, mean, .before = 4, .after = 3))
```

We can plot them individually, or plot all versions on the same plot.

```{r}
aus_jap_arrivals |> select(-Origin) |>
  pivot_longer(cols = contains("Arrivals"), names_to = "Type") |>
  autoplot(value) + scale_color_viridis(discrete = TRUE)
```

There's a bit of overplotting here, but we generally see that the time series 
becomes smoother as we increase the window size. 
This is similar to bandwidth selection for kernel smoothing. 
On the other hand, MA4 seems to be smoother than MA6 (see the figure below). 
This is because the seasonality has period 4, which is averaged out when the
window size of the moving average is a multiple of 4, but not otherwise.

```{r}
aus_jap_arrivals |> select(-Origin) |>
  pivot_longer(cols = all_of(c("Arrivals", "ArrivalsMA4", "ArrivalsMA6")), 
               names_to = "Type") |>
  autoplot(value) + scale_color_viridis(discrete = TRUE)
```

## Log transform and differencing

Let's consider the turnover for clothes sales in Victoria, Australia.

```{r}
diabetes <- read_rds("../_data/cleaned/diabetes.rds")
diabetes |> autoplot()
```

The seasonal fluctuations grow with the level of the time series, so we shall 
apply a log transformation.

```{r}
diabetes |> 
  mutate(logTotalC = log(TotalC)) |>
  autoplot(logTotalC)
```

This is much better now. Let us now take differences

```{r}
diabetes |> 
  mutate(logTotalC = log(TotalC)) |>
  mutate(logDiffTotalC = difference(logTotalC)) |>
  autoplot(logDiffTotalC)
```

This measures the percentage *month-on-month* increase in sales of diabetes drugs. 
The time series still seems to be quite seasonal.

What happens if we take a seasonal difference?

```{r}
diabetes |> 
  mutate(logTotalC = log(TotalC)) |>
  mutate(logSeasonalDiffTotalC = difference(logTotalC, 12)) |>
  autoplot(logSeasonalDiffTotalC)
```

It measures the *year-on-year* percentage increase in sales of diabetes drugs. 
The seasonality has disappeared! This series is also much more stable, see the 
y-axis limits.