---
title: "Week 6 Demonstration"
format: html
editor: visual
---

## Set up

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
library(slider)
library(gridExtra)
```

## 1. ACVF, ACF, and stationarity

Consider the time series defined as $$
X_t = U_1 \sin(2\pi \omega t) + U_2\cos(2\pi\omega t),
$$ where $U_1$ and $U_2$ are independent random variables with zero mean and variance 1.

a.  Compute the ACVF of $(X_t)$.
b.  Is $(X_t)$ weakly stationary?
c.  Is $(X_t)$ strictly stationary? If not, what conditions can we put on $U_1$ and $U_2$ to make it stationary?

## 2. Moving averages and differences

Consider the linear trend model $$
X_t = \beta_0 + \beta_1 t + W_t.
$$ Define a time series $(Y_t)$ by taking a moving average of window 7. Define another times series $(Z_t)$ by taking a difference.

a.  What is the mean function for $(Y_t)$? What is the ACVF for $(Y_t)$?
b.  What is the mean function for $(Z_t)$? What is its ACVF?
c.  What is the CCF of $(Y_t)$ and $(Z_t)$?

## 3. Forecasting

Let $(X_t)$ be a zero mean stationary process with $\rho_X(h) \neq 0$ for some $h$

a.  Find $A$ minimizing $\mathbb{E}\lbrace (X_{n+h} - A X_n)^2 \rbrace$.
b.  Use this to make a forecast for $X_{n+h}$ given $X_1,\ldots,X_n$.
c.  What is the MSE for the forecast?
d.  What is the consequence of this for ACF values?

## 4.Residual analysis

a.  What is the difference between innovation residuals and regular residuals? Why do we perform residual analysis on the former rather than the latter?
b.  Starting with the following code snippet, fit several models to the Australian takeaway turnover time series and analyze their residuals. Which models have a good fit? Compare this with their CV error.

```{r}
takeaway <- aus_retail |>
  filter(Industry == "Takeaway food services") |>
  summarise(Turnover = sum(Turnover))
```
