---
title: "Week 7: Review"
format: html
editor: visual
---

## Set up

```{r}
#| message: FALSE
library(fpp3)
library(tidyverse)
library(slider)
library(gridExtra)
```

## Summary

### Data wrangling and visualization

-   How to create tsibbles?
-   What are the components of the tsibble data strucutre, in particular its index and key variables?
-   How to detect and fill in missing or duplicated entries?
-   How to filter a tsibble?
-   How to make and interpret the following types of plots?
    -   Time plot
    -   Seasonal plot
    -   Seasonal subseries plot
    -   Scatter plots
    -   Lag plots
-   Time series patterns: Trend, seasonality, cycles
    -   How are these defined?
    -   How to recognize them from the plots?

### Transformations

-   How to implement the following transformations in R, when to use them, how to interpret the transformed time series?
    -   Lags
    -   Differences
    -   Seasonal differences
    -   Log transformation
    -   Box-Cox transformations
    -   Aggregation and downsampling
    -   Moving averages and other rolling window transformations
    -   Calendar adjustment
    -   Population adjustment
    -   Inflation adjustment
-   What is the effect of window size on moving averages, especially for seasonal time series?

### Decomposition

-   Time series components: Trend, seasonal and remainder
    -   How are these defined?
    -   How are they related to time series patterns?
-   Classical decomposition, in both additive and multiplicative forms
    -   What are its assumptions?
    -   How is it defined?
    -   How do the assumptions lead to consistent estimation of the components?
    -   How to implement it in R?
-   What are some problems with the classical decomposition, and how can these be overcome?

### Summary statistics

-   What are summary statistics?
-   How are the following summary statistics defined, how can they be computed, what are they used for?
    -   (Sample) mean, variance, quantiles
    -   (Sample) autocovariance and autocorrelation functions
    -   (Sample) cross-correlation
    -   Decomposition statistics: Strength of trend and strength of seasonality
-   ACF plots
    -   How to make and interpret them?
    -   How does the plot look like when the time series is dominated by trend, seasonality, or short-term cycles?
-   What is the use of summary statistics in time series classification?

### Introduction to forecasting

-   How are the following terms precisely defined?
    -   Forecast
    -   Forecast horizon
    -   Forecast error
-   Simple forecasting methods: How they are defined,
    -   Mean method
    -   Naive method
    -   Seasonal naive method
    -   Linear trend method
    -   Drift method
    -   Seasonal naive method with drift
-   How can decomposition and transformations be used in forecasting?
-   What are the statistical models associated with the simple forecasting methods?
-   Prediction intervals
    -   How are they defined?
    -   How can they be calculated and plotted?
    -   How much should we trust them?
-   Forecast errors
    -   What error metrics are used?
    -   Why do we preferred scaled errors? How can we interpret them?
-   Train-test splits
    -   How is it defined? How to implement it in R?
    -   Why can we not perform random splits?
    -   What is their limitation?
-   Time series cross-validation
    -   How is it defined?
    -   How can it be implemented in R?
    -   How can we interpret the error?
    -   What do the `.step` and `.init` parameters mean? And what is the tradeoff between setting a small value or large value?
    -   How many error terms are used in calculating the total error?
    -   How can we calculate the TSCV error for only $h$-step ahead forecasts?

### Exponential smoothing

-   For the following exponential smoothing methods, how are the methods defined? How can they be implemented in R?
    -   Simple exponential smoothing
    -   Holt linear trend method
    -   Damped trend method
    -   Holt-Winters' methods
-   How are the parameters optimized?
-   What is the meaning of each parameter? How can they be computed using R?
-   When is each method preferred?
-   What is the difference between additive and multiplicative versions of Holt-Winters?
-   What happens when terms in the model are not specified when calling `ETS()`?

### Stationary processes

-   What is a time series model?
-   How are weakly and strictly stationary process defined?
-   How is joint stationarity defined?
-   Mean, autocovariance, autocorrelation, and cross-correlation functions
    -   How are they defined?
    -   How to calculate them for simple time series models?
-   How are the following types of time series models defined?
    -   White noise
    -   Random walk
    -   Signal plus noise
    -   Linear process
    -   Gaussian process
-   Asymptotic normality theorems for mean and ACF
    -   What are the assumptions for the two theorems?
    -   What is the limiting behavior for white noise?
-   What is the Box-Jenkins method? What is its relationship to stationary processes?
-   What are two methods of transforming a time series to make it stationary?
-   What are residuals?
-   How can we use an ACF plot to test whether a time series is drawn from a white noise model?
-   Ljung-Box test
    -   What is the test statistic?
    -   What is the limiting behavior of the test statistic?
    -   How is the test defined?
    -   How can we implement it using R?
    -   How can we choose the lag for the test?
-   How to simulate a time series from a time series model in R?

## Questions

a.  Load the data set `sg-births.csv`.

b.  Make time, seasonal, and seasonal subseries plots.

c.  Comment on all interesting features of the plots.

d.  Without using software, sketch an ACF plot of the time series (the numbers don't have to be exactly correct, just the overall pattern).

e.  Will a Box-Cox transformation be appropriate for this time series?

f.  What sort of transformation may help to improve forecasting?

g.  Will a classical decomposition be appropriate for this time series? What could you use to improve on classical decomposition?

h.  Is this time series stationary?

i.  Is the remainder component stationary?

j.  Which of forecasting methods we have learnt may be appropriate for this time series?

k.  Fit an additive Holt-Winters model. How can the model parameters be extracted? How can we visualize the different components of the model? Compute a forecast for March 2024 by hand. Check whether it agrees with the computed value.

l.  Perform time series cross-validation to compare the different models for the forecast horizon $h=3$. Which model would you pick?

m.  Are you satisfied with the best model? Why or why not?

## Solutions

```{r}
sg_births <- read_rds("../_data/cleaned/sg-births.rds")
```

```{r}
sg_births |> autoplot(Births)
```

```{r}
sg_births |> gg_season(Births)
```

```{r}
sg_births |> gg_subseries(Births)
```

```{r}
sg_births |> ACF(Births) |> autoplot()
```

A Box-Cox transformation is not appropriate. It requires the amplitude of the fluctuations of the time series to increase with its level. That does not happen here because .

```{r}

```